Searching for long-period transiting planets in the \kepler\ light curves using supervised classification

### 2015-03-30 document/ms.tex
words about injections, recall, and precision

\citep{kois}

@@

The Random Forest (RF) classification model \citep{Breiman:2001} has become
very popular in the machine learning community.
It has also been applied with great success in astronomy \citep[for
example]{Richards:2011, Richards:2012} \dfmtodo{others}.
The RF model works by building

@@

@@

@@

\section{Tuning parameters}

@@

\section{Demonstration}

>>>>

@@

The Random Forest (RF) classification model \citep{Breiman:2001} has become
very popular in the machine learning community.
It has also been applied with great success in astronomy \citep[for
example]{Richards:2011, Richards:2012} \dfmtodo{others}.
The RF model works by fitting an ensemble of decision trees to randomly
selected subsamples of the training dataset.
Each tree in the forest is ``grown'' by greedily choosing decision boundaries
that optimize the separation of the training data in randomly selected
subsets of the features until the separation is complete, each leaf contains
only one class or, at most, a fixed number of samples.
At each branch, the decision boundary is set by maximizing either the entropy
or the ``gini'' coefficient.

For the problem of single transit search, our problem can be phrased as a
classification problem where we want to assign every $K$ data point long chunk
of light curve into either the \texttt{transit} or \texttt{no transit} class.
This problem doesn't fall under the standard format of a supervised
classification problem because very few long-period transits have actually
been observed or classified.
We do think, however, that we have a good physical generative model for the
signal of interest and most of the observations of each star have no transits.
Therefore, we can train the model on simulated signals.

@@

To search for single transits in the light curve of a specific star, we fit
three random forest models.

@@

We need to train the classification model on simulated signals.
To simulate these training and validation examples, we randomly select
sections of 121 contiguous flux measurements and simulate the transit signal
induced by the physical orbit of a large planet.
The specific distribution of simulated parameters is listed in \dfmtodo{some
table}.
In the simulations, we take limb darkening and integration time into account
\dfmtodo{cite}.
Since we expect most of the systematic variability to be time-reversible, we
augment the training set with simulations injected into the reversed time
series \dfmtodo{cite some examples of rotational invariance}.
The training set also includes an equal number of negative examples---time
series chunks of the same length without a simulated transit injection.
By virtue of the fact that we are specifically searching for single transits
it is safe to assume that none of the negative samples from the other light
curve sections will have no transits.

We also generate a validation set of simulated signals.
This set is identical to the training set except it is simulated in sections
of light curve that are not used for training.
As discussed below, this set is used to determine the transit significance
threshold required to achieve a pure sample.

In practice, we only need to train three models on the three disjoint
training sets and then for each model, we validate and test on both of the
remaining splits.

\paragraph{Features}

In many supervised classification problems, a sophisticated feature
extraction method is used to improve the performance of the procedure.
For this specific problem, we find excellent performance using the 121
normalized flux values as the features.
The depth of the transit and the noise in the light curve is not irrelevant
to detection so we don't normalize the features to unit variance in the
standard way.
Instead, we normalize each sample (feature vector) by it's empirical median
value and take the logarithm.
Another option for a feature set could be the Fourier transform or wavelet
transform of the light curve section.
Alternatively, the spectrogram of the fluxes in a larger window.
We have experimented briefly with this options finding similar performance.

@@

\section{Tuning parameters}\sectlabel{tuning}

Many choices need to be made in building this classification model.
These choices can be specified by a set of parameters that can be tuned the
optimize the performance of the method.
At this point, it's probably worth discussing how one might specifically
measure the ``performance'' of a search for transits.
The goal is, of course, to \emph{find transiting planets}.
Therefore, we would like to develop a method that finds the most planets
possible.
This will, in general, trade off with the purity of the sample.
For example, a method that returns every possible point the light curve as a
planet candidate will be guaranteed to find every signal in the dataset---it
will have excellent completeness or \emph{recall}---but almost every
candidate will be false---the \emph{precision} is low.
This is obviously not what we want.
Instead, the more relevant metric is probably the recall at fixed (high)
precision.

Let's be clear about what we mean by the terms \emph{precision} and
\emph{recall} in this context because they have specific definitions in the
machine learning that will be useful in our discussion here.
Recall is the probability, integrated across the full parameter space of
interest that a true transit signal will be detected by the method.
In astronomy, this quantity is generally called the completeness or the
detection efficiency and it is routinely measured for transit surveys
\dfmtodo{cite}.
On the other hand, precision is the probability that a positive classification
will be true.
In other words, for a given candidate that our pipeline spits out, the
probability that it is a false alarm is given by one minus the precision.
In astronomy, precision is generally hard or impossible to measure because we
have no ground truth; we don't know that \emph{there is definitely no transit}
at a given time in the light curve.
However, with a few weak assumptions, we can, in practice, estimate the
precision of our single transit search robustly.

\paragraph{Estimating the survey precision}

There are two key assumptions that enter into our precision calculation.
The first is that all transiting planets with more than one transit and a
large signal-to-noise ratio have been previously discovered by one of the many
transit searches applied the \kepler\ dataset \dfmtodo{cite}.
This is a safe assumption because, for large planets, these surveys are all
largely complete out to periods yielding only 2 or 3 transits in the \kepler\
baseline \dfmtodo{cite}.
The second necessary assumption is less robust but it is required for a
completely reliable measurement of the survey precision.
This assumption is that the noise processes in the data are stationary.
In other words, the transit-like noise signals in the training and validation
sets must be ``similar'' to the signals in the target section.
The PDC processing of the data successfully removes most of the
instrument-induced signals with some exceptions that are discussed in
\sect{demo}.
Perhaps we make the further assumption that the stellar variability is
also stationary.

Under these assumptions, the precision of the search procedure is
estimated using a set of injections.

@@

\section{Demonstration}\sectlabel{demo}
